%*******************************************************
% Acknowledgments
%*******************************************************
\pdfbookmark[1]{Acknowledgments}{acknowledgments}

\bigskip

\begingroup
\let\clearpage\relax
\let\cleardoublepage\relax
\let\cleardoublepage\relax
\chapter*{Acknowledgments}

First of all, I want to thank my supervisor Prof. Dr.-Ing. Bernd Edler for all his help and support throughout all the years. Bernd is the nicest and kindest supervisor anyone could have: his door was always open for me to have fruitful discussions and solve tricky problems together and at the same time he was giving me enough freedom to allow me to develop my own research ideas.

\bigskip

Besides my advisor, I want to thank Prof. Gaël Richard for taking the time to review my thesis.

\bigskip
I also want to thank Dr. Antoine Liutkus for inviting me to Nancy for a summer research visit that resulted in so many good ideas and a great friendship!

\bigskip

I want to thank all the amazing people in the AudioLabs, which made my work in Erlangen so enjoyable: First, I want to thank Elke, Tracy and Day-See for all the administrative help and beyond. Then, I want to thank Stefan Turowski for his great technical management in the AudioLabs. Next, I want to thank all of the colleagues at the AudioLabs (in alphabetical order): 
Alexander Adami, Stefan Bayer, Stefan Balke, Sebastian Braun, Tom Bäckström, Soumitro Chakrabarty, Youssef El Baba, Alexandra Craciun, Christian Dittmar, Sascha Disch, Jonathan Driedger, Esther Feichtner, Johannes Fischer, Yesenia Lacouture Parodi, Emanuël Habets, Jürgen Herre, Nanzhu Jiang, Patricio Lopez-Serrano, Wolfgang Mack, Goran Markovic, Vlora Arifi Müller, Meinard Müller, Thomas Prätzlich, Sebastian Rosenzweig, Konstantin Schmidt, Michael Schöffler, Armin Taghipour, Stefan Turowski, Maja Taseska, Maria Luis Valero, Elke Weiland, Christof Weiß, Nils Werner, Frank Zalkow and Julia Zalkow. Thanks to the people at Fraunhofer IIS, especially to Sascha Disch, Christopher Oates, Frederik Nagel, Christian Uhle.
And I also want to thank Thomas Zeiser from the RRZE high performance cluster for his great support.
In this vein, I want thank to all the great scientific open source tools out there that powered most of the experiments in this thesis.
\bigskip

A big thank to the students and interns I supervised and I thank them for all the great work: Berkan Ercan, Erik Johnson, Aravindh Krishnamoorty, Jeremy Hunt, Bufei Liu and Qiao Wang. 
To Karlheinz Busch from the Bamberg Symphonic Orchestra, Johannes Huber and Michael Müller for making our datasets possible. 

\bigskip

To Annika, Lisa, Florian, Chris and Chris for the great time Nuremberg and to my great from Hamburg: Johannes, Malte, Makoto, Robin, Mathis. To Mathieu, Cheryl and Elias for the warm welcome in Montpellier.
I also want to thank the Faller family for their great support during the hard times of writing this thesis.

\bigskip

I deeply thank my family --- Dagmar, Heinrich and Marion --- for supporting me in every point of time in my life.

\bigskip

And last but not least, I want to thank my beloved partner and friend Claudia who gave me so much joy and hope that this journey succeeds.

\bigskip

...and to my son Emil for his beautiful smile.

\endgroup
